% ===============================================================
% HODGKIN-HUXLEY: RIGOROUS MATHEMATICAL FORMULATION
% ===============================================================

\section{Refined Hodgkin--Huxley Model with Corrected Sodium Inactivation}

\subsection{State Space, Domain, and Units}

The state vector is
\[
(V, m, h, n) \in \mathbb{R} \times [0,1]^3,
\]
where $V$ is the membrane voltage (mV), and $m,h,n$ are gating probabilities.
The system follows the Hodgkin--Huxley formalism \cite{hodgkin1952quantitative}.

All parameters use the standard squid-axon convention:
\[
C_m \, [\mu\mathrm{F}/\mathrm{cm}^2], \quad
\bar g \, [\mathrm{mS/cm}^2], \quad
I \, [\mu\mathrm{A/cm}^2], \quad
\alpha_x,\beta_x \, [\mathrm{ms}^{-1}].
\]

\subsection{Membrane Potential Equation}

The membrane equation is given by conservation of charge:
\begin{equation}
\label{eq:dVdt}
C_m \frac{dV}{dt}
= I_{\mathrm{stim}}(t)
- \left[
\bar g_{\mathrm{Na}} m^3 h (V - E_{\mathrm{Na}})
+ \bar g_{\mathrm{K}} n^4 (V - E_{\mathrm{K}})
+ \bar g_{\mathrm{L}} (V - E_{\mathrm{L}})
\right].
\end{equation}

\subsection{Gating Variable Dynamics}

Each gating variable $x \in \{m,h,n\}$ satisfies:
\begin{equation}
\label{eq:dxdt}
\frac{dx}{dt} = \alpha_x(V) \left( 1 - x \right) - \beta_x(V) x.
\end{equation}

Steady state:
\[
x_\infty(V) = \frac{\alpha_x(V)}{\alpha_x(V) + \beta_x(V)}.
\]


\subsection{Kinetic Rate Functions \texorpdfstring{$\alpha_x, \beta_x$}{alphabeta}}

Table \ref{tab:rates} lists the refined transition rates.  
The sodium inactivation gate $h$ includes the corrected formulation verified in our depolarization block benchmark.

\begin{table}[h!]
\centering
\caption{Hodgkin--Huxley transition rate functions (with corrected $h$-gate).}
\label{tab:rates}
\begin{tabular}{lll}
\hline
Variable & $\alpha_x(V)$ & $\beta_x(V)$ \\
\hline
$m$ (Na activation) &
$\displaystyle
\frac{0.1 (25 - V)}{e^{(25-V)/10}-1}
$ &
$\displaystyle
4 e^{-V/18}
$ \\[1.2em]
$h$ (Na inactivation) \textbf{(Corrected)} &
$\displaystyle
\frac{1}{e^{(30-V)/10}+1}
$ &
$\displaystyle
0.07 e^{-V/20}
$ \\[1.2em]
$n$ (K activation) &
$\displaystyle
\frac{0.01 (10 - V)}{e^{(10-V)/10}-1}
$ &
$\displaystyle
0.125\, e^{-V/80}
$ \\
\hline
\end{tabular}
\end{table}

Singularities at $V=25$ or $V=10$ are resolved by using analytical limits.

\subsection{Existence, Uniqueness, and Invariance}

The right-hand side of the system $(\ref{eq:dVdt})$--$(\ref{eq:dxdt})$ is locally Lipschitz on
$\mathbb{R} \times [0,1]^3$.
Thus, by the Picard--Lindelöf theorem,
a unique local solution exists for all initial conditions.

Positivity invariance holds:
\[
x(0)\in[0,1]\quad \Rightarrow\quad x(t)\in[0,1],\;\forall t\ge 0,
\]
because the vector field points inward at boundaries.

\subsection{Analytical Result: Depolarization Block}

At depolarized resting potential $V_{\mathrm{rest}}=-40\,\mathrm{mV}$, the corrected inactivation gate fulfills:
\[
h_\infty(-40) \approx 0.0018.
\]

Thus,
\[
m^3 h \approx 0,
\]
preventing the sodium current from generating an action potential.  
Therefore, stimulus-induced spike initiation fails:
\[
\max_t V(t) \approx -54.4\,\mathrm{mV}.
\]
This reproduces the classical depolarization block phenomenon observed in channelopathy and high-frequency failures.

